# Data transformation
(with the feature details used the future analysis that might be confused)


## Social Vulnerability Index 
E_TOTPOP: Population estimate

E_POV: Persons below poverty estimate

EP_POV: Percentage of persons below poverty estimate

E_PCI: Per capita income estimate

E_NOHSDP: Percentage of persons with no high school diploma (age 25+)

E_UNEMP: Civilian (age 16+) unemployed estimate

EP_UNEMP: Unemployment Rate estimate
```{r}
# length(sv$ST) # 3142
# sum(is.na(sv))
# unique(sv[c("STATE", "COUNTY")]) # 3142
sv2 <- sv %>% select(ST, STATE, COUNTY, E_TOTPOP, E_POV, EP_POV, E_PCI, 
                     E_NOHSDP, EP_AGE65, EP_AGE17, EP_MINRTY, E_UNEMP, 
                     EP_UNEMP) %>%
  mutate(EP_NOHSDP = round(E_NOHSDP/E_TOTPOP*100, 2)) %>%
  filter(EP_POV >=0)
# 
head(sv2)

```

## COVID-19 Deaths
Population: Total population among the corresponding region

Deaths: Number of COVID-19 deaths 
```{r}
cov19.death <- cov19_death %>% dplyr::select(-c(`Notes`, `Census Region Code`, `State Code`, `County Code`, `Race Code`, `Crude Rate`))
# cov19.death
# unique(cov19.death$Race)
# cov19.death %>% filter(is.na(`Census Region`))
cov19.death <- cov19.death %>% filter(!is.na(`Census Region`))
head(cov19.death)

```

## 2021 COVID-19 Vaccine Attitudes among People Experiencing Homelessness in Los Angeles, CA
phq4_dum: The Four-Item Patient Health Questionnaire for Anxiety and Depression

phq4_sum: PHQ4 sum score

fear_sum2: COVID-19 threat index (sum score of wb011, wb012, wb014, wb015), indicating the degree of fear toward COVID-19

sd_sum3: COVID-19 protective behavior (sum score of handwash, sd01_sixfeet, mask_always) 

vacc_hes: Hesitancy toward receiving COVID-19 vaccine? 

vacc_if: If offered COVID-19 vaccine, would you take it?
```{r}
la <- la[-1, ]
la2 <- la%>% select(age, female, race2, housing2, years_homeless, phq4_dum, 
                      phq4_sum, fear_sum2, sd_sum3, at02_official_any, 
                      at02__newspapersradioortv, at02_ff_any, vacc_hes,
                      vacc_offered, vacc_if)
col.names <- c("Age", "Gender", "Race", "Housing")
colnames(la2)[1:4] <- col.names

hes.reason <- la %>% select(reason_risk, reason_protection, 
                            reason_information, reason_sideeffect,
                            reason_moreppl, reason_novax, reason_riskgroup, 
                            reason_afford) %>% drop_na()

head(la2)
hes.reason
```

## Geographic Data
```{r}
head(shape)
```

## Vaccination by Race 
key: percent of vaccinated among race in correspoding region
```{r}
colnames(vaccine_by_race)[1] <- "state"
colnames(vaccine_by_race)[2] <- "percent of white vaccinted"
colnames(vaccine_by_race)[3] <- "percent of black vaccinted"
colnames(vaccine_by_race)[4] <- "percent of Hispanic vaccinted"
colnames(vaccine_by_race)[5] <- "percent of Asian vaccinted"


vaccine_by_race[(vaccine_by_race)== "NR"] <- NA

# colSums(is.na(vaccine_by_race)) %>% sort(decreasing = TRUE)
# vaccine_by_race %>% filter(is.na(`percent of Hispanic vaccinted`)) %>% pull(state)
head(vaccine_by_race)
```